---
output:
  html_document:
    code_folding: hide
---

# Metadata for Grasslands Data Package {.tabset}
**`r format(Sys.time(), '%B %d, %Y')`**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary
This script creates EML metadata for Southern Plains Network Grasslands data packages in DataStore. The script must be run separately for internal and public versions of the data package. Records for the data packages in DataStore should be created prior to running this script.

The metadata generated will be of sufficient quality for the data package Reference Type and can be used to automatically populate the DataStore fields for this reference type. This is a step by step process where each section should be reviewed, edited if necessary, and run one at a time. Several EMLassemblyline functions are decision points and may only apply to certain data packages. This workflow takes advantage of the NPSdataverse, an R-based ecosystem that includes external EML creation tools such as the R packages EMLassemblyline and EML. However, these tools were not designed to work with DataStore. Therefore, the NPSdataverse and this workflow also incorporate steps from NPS-developed R packages such as EMLeditor and DPchecker. You will necessarily over-write some of the information generated by EMLassemblyline. That is OK and is expected behavior.

This script is based on the EML generation template from NPSdataverse. Melina Kompella initially authored this script based on the template and Cheryl McIntyre and Amy Luo revised these scripts. 

### Good additional references
[EMLassemblyline](https://ediorg.github.io/EMLassemblyline/)
[EMLeditor](https://nationalparkservice.github.io/EMLeditor/index.html)

## Clone DataStore Record for Prior Version

Prior to running this script, clone the DataStore record for the most recently published version of the grasslands data for SOPN. Change the title to something that makes it obvious that it's a draft. The title will be updated when the metadata are extracted later. In the .Renviron, populate datastore_ref with the newly generated DataStore record/reference number. Save the new record as a draft.

## Install and Load R Packages 

Install packages. If you have not recently installed packages, please re-install them. You will want to make sure you have the latest versions of all the NPSdataverse packages - QCkit, EMLeditor, DPchecker, and NPSutils - as they are under constant development. If you run into errors installing packages from github on NPS computers you may first need to run:

```{r renviron, eval=FALSE}
options(download.file.method="wininet")
```
```{r package_install, eval=FALSE}
#install packages
install.packages(c("devtools", "tidyverse"))
devtools::install_github("nationalparkservice/NPSdataverse")
```
```{r load_packages}
# When loading packages, you may be advised to update to more recent versions
# of dependent packages. Most of these updates likely are not critical.  
# However, it is important that you update to the latest versions of QCkit, 
# EMLeditor, DPchecker and NPSutils as these NPS packages are under constant 
# development.
library(NPSdataverse)
library(tidyverse)
```

## Set the overall package details

All of the following items should be reviewed and updated to fit the package you are working on. For lists of more than one item, keep the same order (i.e. item #1 should correspond to the same file in each list).

### Set values in .Renviron

The script relies on values in .Renviron, which you may need to update:

- start_year: the beginning year of data in your analysis (or data package)
- end_year: the last year of data in your analysis (or data package)
- dbExports: where the exports from the Grasslands database are stored
- dbExportDate: the date of the dbExports from the Grasslands database, which should be at the end of each file name. Format as YYYYMMDD, e.g., 20230802.
- EXPORT_INTERNAL_DATAPACKAGE: file path where to save the data package files, if internal
- EXPORT_PUBLIC_DATAPACKAGE: file path where to save the data package files, if public
- package_version: whether data package is public or internal

#### Set internal or public data package

```{r package-version}
if (package_version == "internal") {
  EXPORT_DATAPACKAGE <- EXPORT_INTERNAL_DATAPACKAGE
} else {
  EXPORT_DATAPACKAGE <- EXPORT_PUBLIC_DATAPACKAGE
}
```

#### Metadata filename

The code chunk below will generate the metadata file name based on the start_year, end_year, and package_version. It ends in _metadata to comply with data package specifications and the extension (.xml) does not need to be included.

```{r metadata_id}
metadata_id <- paste("SOPN_Grassland", start_year, end_year, package_version, "metadata", sep = "_")
```

#### Package Title

The code chunk below will generate the data package title name based on the start_year, end_year, and package_version. It complies with FAIR principles, which suggest titles of between 7 and 20 words.

```{r package_title}
package_title <- paste0("Grassland Vegetation Monitoring Data for the Southern Plains Network (SOPN): ", start_year, "-", end_year, " Data Package (", str_to_upper(package_version), ")")
```

#### Data collection status

Set data_type to "ongoing" (rather than "complete", which implies that we will not collect any more data).

```{r data_status}
data_type <- "ongoing"
```

#### Path to data file(s)

Tell R where your data files are. The file path is defined by the EXPORT_DATAPACKAGE variable in your .Renviron. Do _not_ push your username to GitHub! 

The data files should _NOT_ be in the working directory or pushed to GitHub. 

The scripts will create a subfolder (TEMP_4eml) within your EXPORT_DATAPACKAGE directory for temporary storage of the files required to generate the metadata.

```{r create-output-dir}
if (file.exists(paste0(EXPORT_DATAPACKAGE, "TEMP_4eml/"))) {
 cat("The folder already exists")
} else {
 dir.create(paste0(EXPORT_DATAPACKAGE, "TEMP_4eml/"))
}

```

#### List data files

Tell R what your data files are called. 

```{r list_files}
#only .csv files in the new directory
data_files <- list.files(EXPORT_DATAPACKAGE, pattern = "*.csv", ignore.case = TRUE)

```

#### Name the data files

These names are relatively short, but perhaps more informative than the actual file names. Make sure that they are in the same order as the files in data_files (should only be an issue if filenames change).

```{r data_names}
data_names <- c("Canopy Closure",
                "Vegetation Presence and Cover & Disturbance History",
                "Fire History",
                "Vegetation Cover, Soil Substrate Cover & Disturbance History",
                "Plant Species List",
                "Transect Origin and End Points",
                "Visit Information",
                "Protocol and SOP Information")
```

#### Describe each data file

Data file descriptions are unique and about 10 words long. Descriptions will be used in auto-generated tables on DataStore. Again, make sure these are in the same order as the files they are describing.

```{r data_descriptions}
data_descriptions <- c("Canopy closure, measured via spherical densiometer",
                       "Vegetation presence in subplots; disturbance history included. 2022-23 data include cover estimates",
                       "Fire history (prescribed burns and wildfires) of grassland plots",
                       "Vegetation and soil substrate cover along transects; disturbance history included",
                       "Plants species list and associated attributes",
                       "Coordinates of rebar monuments marking plot transects",
                       "Visit (sampling event) information",
                       "SOP version history and links to DataStore references")
```

#### Placeholder URL for data files

EMLassemblyline needs to know where the data files will be (a URL). For now, insert a place holder. Don't worry--this information will be updated later on when you add a Digital Object Identifier (DOI) to the metadata.

```{r temp_data_url}
data_urls <- c(rep("temporary URL", length(data_files)))
```

#### Taxonomic information

Specify where your taxonomic information is. In this case, the _PlantSpecies file compiles all of the species so we can just point to it and the scientificName field within it. 

```{r taxonomic_info}
# the file(s) where scientific names are located:
data_taxa_tables <- "SOPN_Grassland_PlantSpecies.csv"

# the column where your scientific names are within the data files.
data_taxa_fields <- "scientificName"
```

#### Geographic information

Specify the table and fields that contain geographic coordinates and site names. This information will be used to fill out the geographic coverage elements of the metadata.

```{r geo_info}
data_coordinates_table <- "SOPN_Grassland_PlotLocations.csv"
data_latitude <- "decimalLatitude"
data_longitude <- "decimalLongitude"
data_sitename <- "locationID"
```

#### Temporal information

Use the Visit Information data table file to get the collection date of the first and last data point in the data package. The YYYY-MM-DD format complies with the International Standards Organization's standard 8601 and should not cause errors down the road. Setting the date to the future will cause errors down the road!

```{r temp_info}
Visits <- read.csv(paste0(EXPORT_DATAPACKAGE,"SOPN_Grassland_PlotVisits.csv")) %>%
  mutate(eventDate = as.Date(eventDate, "%Y-%m-%d"))

startdate <- lubridate::ymd(min(Visits$eventDate))
enddate <- lubridate::ymd(max(Visits$eventDate))

rm(Visits)
```

## EMLassemblyline Templates

The next set of code and functions will copy the existing metadata templates from the grasslands-r repo into the TEMP_4eml subfolder within your EXPORT_DATAPACKAGE folder. It will also generate the geographic and taxonomic component, which are not included in the templates or pushed to GitHub to prevent inadvertent release of information.

#### Core Metadata Information

This code chunk will _copy_ the metadata template files from the grasslands-r folder to your TEMP_4eml subfolder within your EXPORT_DATAPACKAGE. You should not need to update most (or any) of these files. Typically these files can be reused between years. But if you need to edit the the metadata template files, it is best to do so in in Excel as they have columns. 

```{r template-copy}
metadata_templates <- list.files("./metadata_templates/",pattern = "*.txt", ignore.case = TRUE)
file.copy(from=paste0("./metadata_templates/", metadata_templates),to=paste0(EXPORT_DATAPACKAGE, "TEMP_4eml/", metadata_templates), overwrite = TRUE, recursive=FALSE)
```

Note that in the personnel file, we list at least one person with a "creator" role and one person with a "contact" role (the same person can be listed for both). Creators are considered authors and will be included in the data package citation. We do not need to list anyone as the primary investigator, so you can ignore the warning about the "PI" being missing. You can also list as many additional people with custom roles as desired (e.g. "field technician", "laboratory assistant"). Each individual must have a surName. The electronicMailAddress is an email and the userId is the person's [ORCID](https://orcid.org) (if they have one). List the ORCID as just the 16 digit string; do not include the https://orcid.org prefix (e.g. xxxx-xxxx-xxxx-xxxx). You can leave the projectTitle, fundingAgency, and fundingNumber blank.

#### Geographic Coverage

This function creates a geographic_coverage.txt file that lists your sites as points based on the lat/long in the Plot Locations table. The geographic_coverage.txt file is stored in your TEMP_4eml and should not be pushed to GitHub.

```{r geo_cov}
EMLassemblyline::template_geographic_coverage(
  path = paste0(EXPORT_DATAPACKAGE, "TEMP_4eml"),
  data.path = EXPORT_DATAPACKAGE,
  data.table = data_coordinates_table,
  lat.col = data_latitude,
  lon.col = data_longitude,
  site.col = data_sitename,
  write.file = TRUE
  )
```

#### Taxonomic Coverage

This function creates a taxonomic_coverage.txt file based on the Plant Species table. The taxonomic_coverage.txt file is stored in your TEMP_4eml and should not be pushed to GitHub.

Currently supported authorities are 3 = ITIS, 9 = WORMS, and 11 = GBIF. In the example below, the function will first try to find the scientific name at ITIS and if it fails will then look at GBIF. If you have a lot of taxa, this could take some time to complete. 

```{r tax_cov}
EMLassemblyline::template_taxonomic_coverage(
  path = paste0(EXPORT_DATAPACKAGE, "TEMP_4eml"), 
  data.path = EXPORT_DATAPACKAGE, 
  taxa.table = data_taxa_tables,
  taxa.col = data_taxa_fields, 
  taxa.authority = c(3,11),
  taxa.name.type = 'scientific', 
  write.file = TRUE
  )
```

## Create an EML File

Run this (it may take a little while) and see if it validates (you should see 'Validation passed'). It will generate an R object called "my_metadata". The function may alert you of some issues that you need to review. Run the function `issues()` at the end of the process to get feedback on items that might be missing or need attention. Fix these issues and then re-run the `make_eml()` function.

```{r make_eml}
my_metadata <- EMLassemblyline::make_eml(
  path = paste0(EXPORT_DATAPACKAGE, "TEMP_4eml"),
  dataset.title = package_title,
  eml.path = EXPORT_DATAPACKAGE,
  data.path = EXPORT_DATAPACKAGE,
  data.table = data_files,
  data.table.name = data_names,
  data.table.description = data_descriptions,
  data.table.url = data_urls,
  temporal.coverage = c(startdate, enddate),
  maintenance.description = data_type,
  package.id = metadata_id,
  return.obj = TRUE, 
  write.file = FALSE
  )
```

## Check for EML validity 

This is a good point to pause and test whether your EML is valid. 

```{r validate}
EML::eml_validate(my_metadata)
```

If your EML is valid you should see the following (admittedly cryptic) result:

```
# [1] TRUE
# attr(,"errors")
# character(0)
```

If your EML is not schema valid, the function will notify you of specific problems you need to address. We HIGHLY recommend that you use the EMLassemblyline and/or EMLeditor functions to fix your EML and do not attempt to edit it by hand.

## Add NPS specific fields to EML 

Now that you have valid EML metadata, you need to add NPS-specific elements and fields. For instance, unit connections, DOIs, referencing a DRR, etc. More information about these functions can be found at: [https://nationalparkservice.github.io/EMLeditor/](https://nationalparkservice.github.io/EMLeditor/).

#### Add Controlled Unclassified Information (CUI) codes

This is a required step. It is important to indicate not only that your data 
package contains CUI, but also to inform users if your data package does NOT
contain CUI because empty fields can be ambiguous (does it not contain CUI or
did the creators just miss that step?). You can choose from one of five CUI
dissemination codes. Watch out for the spaces! These are:

  * PUBLIC - Does NOT contain CUI.
  * FED ONLY - Contains CUI. Only federal employees should have access (similar to "internal only" in DataStore).
  * FEDCON - Contains CUI. Only federal employees and federal contractors should have access (also very much like current "internal only" setting in DataStore).
  * DL ONLY - Contains CUI. Should only be available to a named list of individuals (where and how to list those individuals TBD)
  * NOCON - Contains CUI. Federal, state, local, or tribal employees may have access, but contractors cannot.
More information about these codes can be found at: [https://www.archives.gov/cui/registry/limited-dissemination](https://www.archives.gov/cui/registry/limited-dissemination)

```{r cui_dissemination}
my_metadata2 <- EMLeditor::set_cui_code(my_metadata, "FEDCON")
# note that in this case I have added the CUI code to the original R object, 
# "my_metadata" but by giving it a new name, i.e. "my_meta2" I could have
# created a new R object. Sometimes creating a new R object is preferable 
# because if you make a mistake you don't need to start over again.
```

## Intellectual Rights

EMLassemblyine and ezEML provide some attractive looking boilerplate for setting the intellectual rights. It looks reasonable and so is easy to just keep. However, NPS has some specific regulations about what can and cannot be in the intellectualRights tag. Use `set_int_rights()` to replace the text with NPS-approved text. Note: You must first add the CUI dissemination code using `set_cui_code()` above, as the dissemination code and license must agree. That is, you cannot give a data package with a PUBLIC dissemination code a "restricted" license (and vice versa: a restricted data package that contains CUI cannot have a public domain or CC0 license). You can choose from one of three options, but for most public NPS data packages, the CC0 license is appropriate:

  * "restricted": If the data contains Controlled Unclassified Information (CUI), the intellectual rights must read: **"This product has been determined to contain Controlled Unclassified Information (CUI) by the National Park Service, and is intended for internal use only. It is not published under an open license. Unauthorized access, use, and distribution are prohibited."**

  * "CC0": If you need a license, for instance if you are working with a partner organization that requires a license, use CC0: **"The person who associated a work with this deed has dedicated the work to the public domain by waiving all of his or her rights to the work worldwide under copyright law, including all related and neighboring rights, to the extent allowed by law. You can copy, modify, distribute and perform the work, even for commercial purposes, all without asking permission."**

  * "public": If the data do not contain CUI, the default is the public domain. The intellectual rights must read: **"This work is in the public domain. There is no copyright or license."**

The `set_int_rights()` function will also put the name of your license in a field in EML for DataStore harvesting.

```{r int_rights}
# choose from "restricted", "public" or "CC0" (zero), see above:
my_metadata2 <- EMLeditor::set_int_rights(my_metadata2, "restricted")
```

#### Add data package reference ID and DOI 

Since you already have a draft reference on DataStore that you want to upload your data package to, you can add the DOI to metadata without creating a new draft reference using the code chunk below where datastore_ref corresponds to the 7-digit DataStore reference ID.

```{r DS_manual_doi}
my_metadata2 <- EMLeditor::set_doi(my_metadata2, datastore_ref)
my_metadata2 <- EMLeditor::set_data_urls(my_metadata2)
```

#### Set the language

This is the human language (as opposed to computer language) that the data package and metadata are constructed in. Examples include English, Spanish, Navajo, etc. A full list of available languages is available from the Library of Congress. Please use "English" as an input.

```{r language}
my_metadata2 <- EMLeditor::set_language(my_metadata2, "English")
```

#### Add content unit links

These are the park units where data were collected from. In this case, we list all of the SOPN parks with grasslands monitoring. The geographic coordinates corresponding to bounding boxes for each park unit listed will automatically be generated and inserted into the metadata. Individual park units are more informative than the bounding box for the entire network.

```{r content_units}
park_units <- c("BEOL", "CAVO", "CHIC", "FOLS", "FOUN", 
                 "LAMR", "LYJO", "PECO", "SAND", "WABA")
my_metadata2 <- EMLeditor::set_content_units(my_metadata2, park_units)
```

#### Add the Producing Unit(s)

Add the unit responsible for generating the data package: SOPN.

```{r prod_units}
my_metadata2 <- EMLeditor::set_producing_units(my_metadata2, "SOPN")

```

#### Add your data package to a DataStore project

Add the SOPN grasslands project to the metadata. Once the metadata are extracted on DataStore, your data package will automatically be added to the DataStore project. Keep in mind that whoever uploads and extracts the metadata on DataStore must have ownership level permissions on both the data package and the project. In addition, you may run into some errors when running the code chunk below if you are not an owner of the SOPN grasslands project. 

```{r add_project}
# where "project_reference", defined in .Renviron, is the DataStore Reference id 
# for the Project that the data package should be linked to.
my_metadata2 <- EMLeditor::set_project(my_metadata2, project_reference)
```

## Validate your EML 

Almost done! This is another great time to validate your EML and make sure
everything is schema valid. Run:

```{r val_eml}
EML::eml_validate(my_metadata2)
```

if your EML is valid you should see the following (admittedly cryptic):

# [1] TRUE
# attr(,"errors")
# character(0)

If your EML is not schema valid, the function will notify you of specific problems you need to address. We HIGHLY recommend that you use the EMLassemblyline and/or EMLeditor functions to fix your EML and do not attempt to edit it by hand.

## Write your EML to an xml file

Now it's time to convert your R object to an .xml file and save it to the directory that you defined in .Renviron.

```{r write_eml}
EML::write_eml(my_metadata2, paste0(EXPORT_DATAPACKAGE, metadata_id, ".xml"))
```


## Check your data package

If your data package is now complete, you can run some tests prior to upload to make sure that the package meets a minimal set of requirements and that the data and metadata are properly specified and consistent with each other. It will run all of the checks in the code chunk directly above, and then some!

```{r congruence_tests}
# This assumes that you have written your metadata to an .xml file and that the .xml file is in the EXPORT_DATAPACKAGE directory with the .csv files.
DPchecker::run_congruence_checks(EXPORT_DATAPACKAGE, skip_cols = c("fireDate", "fieldYear"))

```

## Upload your data package 

If everything checked out, you should be ready to upload your data package! 

1. Within DataStore, open the record that you created and click edit.
2. Navigate to Files and Links and upload the csv and xml files. 
3. Extract all the information from the metadata file to populate the relevant DataStore fields. On the left side of the files list, select the radio button corresponding to your metadata. Then click the "Extract Metadata" button from the top right of the box with the files listed in it. 
4. Carefully review the DataStore record and ensure that the metadata populated correctly. It should but this is also a good time to double-check for errors.
5. Carefully review the permissions tab and double-check that the right people can download files.
6. The public data package should be peer-reviewed before being released. Ask a trusted co-owner to review the record. 
7. Once you’re given the go ahead, take a deep breath and activate it.